---
title: "Introduction to EIVmodels"
description: >
  Learn how to get started
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to EIVmodels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of EIVmodels is to fit models accounting for measurement error. 

## Setup

```{r}
library(EIVmodels)
```

## Example - Simple Linear Regression

This package provides options to simulate data from different types of models where there is measurement error in both the `x` and `y` variable. We'll start by simulating some data from a linear regression model. 

```{r}
dat <- sim_slr(n_sim = 10,
               alpha = 0,
               beta = 1,
               x_err = 0.1)
dat
```

Next we'll plot the simulated data and we'll add the true regression line to the plot by adding the argument `add_truth = TRUE` to the `plot_dat()` function. 

```{r}
plot_dat(dat,
         add_truth = TRUE)
```

Now we'll fit the errors-in-variables simple linear regression model to the data and see how close we get to estimating the true data generating process. We'll start by creating a model object and specifying the argument `model = "model_eiv_reg"`. You'll see some messages and a progress bar pop up as the JAGS model starts to run. 

```{r}
mod_eiv_slr <- run_mod(dat,
                   model = "model_eiv_slr")
```

Now we'll plot the model-based estimates with uncertainty. We'll overlay the true line as a comparison. 

```{r, warning=FALSE, message=FALSE}
plot_res(mod_eiv_slr,
         add_truth = TRUE)
```

To access the data that created this plot, use:

```{r}
mod_reg_res <- par_est(mod = mod_eiv_slr)
mod_reg_res$pred_res
```

To access a summary of the parameter estimates, use:

```{r}
mod_reg_res$par_summary
```

