---
title: "Introduction to EIVmodels"
description: >
  Learn how to get started
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to EIVmodels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

EIVmodels is a R package designed specifically to account for measurement errors within some commonly used models (linear regression, change-point regression, (Integrated) Gaussian process regression) when analysing time-dependent data derived from paleoenvironmental reconstructions.

## Setup

```{r}
library(EIVmodels)
```

## Example - Simple Linear Regression

This package provides options to simulate data from different types of models where there is measurement error in both the `x` and `y` variable. We'll start by simulating some data from a linear regression model with measurement error. Below I'm simulating 20 data points where the intercept ($\alpha$) = 0, slope ($\beta$) = 1. The measurement errors for $y$ and $x$ are set to 0.1. 

```{r}
dat <- sim_slr(n_sim = 20,
               alpha = 0,
               beta = 1,
               y_err = 0.1,
               x_err = 0.1)
dat
```

The output data contains the simulated $x$, $y$ data as well as the measurement errors that were supplied. It will also contain the "true" underlying process, which can be used for comparison purposes later. 

Next we'll plot the simulated data and we'll add the true regression line to the plot by adding the argument `add_truth = TRUE` to the `plot_dat()` function. 

```{r}
plot_dat(dat,
         add_truth = TRUE)
```

Now we'll fit the errors-in-variables simple linear regression model to the data and see how close we get to estimating the true data generating process. We'll start by creating a model object and specifying the argument `model = "slr"`. You'll see some messages and a progress bar pop up as the JAGS model starts to run. 

```{r}
mod_eiv_slr <- run_mod(dat,
                       model = "slr")
```

Now we'll plot the model-based estimates with uncertainty. We'll overlay the "true" line as a comparison. 

```{r, warning=FALSE, message=FALSE}
plot_res(mod_eiv_slr,
         add_truth = TRUE)
```

To access the data that created this plot, use:

```{r}
mod_reg_res <- par_est(mod = mod_eiv_slr)
mod_reg_res$pred_summary
```

To access a summary of the parameter estimates, use:

```{r}
mod_reg_res$par_summary
```

## Example - Using real data

The package contains a dataset called `NJ_CC` which is subset of a sea level reconstruction from New Jersey, USA. Note the format of the dataset. This is the format you should use for any data you want to apply these models to. 

```{r}
EIVmodels::NJ_CC
```
Let's plot the data

```{r}
plot_dat(NJ_CC)
```

__EIV simple linear regression__

Let's fit the EIV simple linear regression model. With these data its a good idea to scale the age data. Dividing my 1000 is useful here, as the relative sea level data ($y$) is in metres, dividing by 1000 means the slope (rate of change) is still easily interpreted in relation to the original scale of the data. In this case the slope can be interpreted in mm/yr.

```{r}
mod_eiv_slr <- run_mod(NJ_CC,
                       model = "slr",
                       scale_factor = 1000)
```

Let's get the parameter estimates

```{r}
mod_reg_res <- par_est(mod = mod_eiv_slr)
mod_reg_res$par_summary
```

Based on this model, the rate of sea level change is 1.9 mm/yr (90% UI: 1.7 to 2.1 mm/yr) 

Now, let's visualise the results

```{r, warning=FALSE, message=FALSE}
plot_res(mod_eiv_slr)
```

__EIV change-point linear regression__

Let's fit the EIV change point linear regression model instead. All we need to do is change the model argument to "cp".

```{r}
mod_eiv_cp <- run_mod(NJ_CC,
                       model = "cp",
                       scale_factor = 1000)
```

Let's get the parameter estimates

```{r}
mod_cp_res <- par_est(mod = mod_eiv_cp)
mod_cp_res$par_summary
```

Now, let's visualise the results

```{r, warning=FALSE, message=FALSE}
plot_res(mod_eiv_cp)
```

__EIV Integrated Gaussian Process__

Let's fit the EIV integrated Gaussian process model. This is a much richer model and will take longer to run than the others. However, an advantage is that it  provides the underlying rate process as well as the sea level process over time. 


```{r, eval = FALSE}
mod_eiv_igp <- run_mod(NJ_CC,
                       model = "igp",
                       scale_factor = 1000)
```

```{r, echo = FALSE, include=FALSE}
mod_eiv_igp <- readRDS("mod_eiv_igp.rds")
```

Let's visualise the results

```{r, warning=FALSE, message=FALSE}
plot_res(mod_eiv_igp)
```


Let's look at the output that creates these results plots

```{r}
mod_igp_res <- par_est(mod = mod_eiv_igp)
mod_igp_res$pred_summary
```

Let's check parameter estimates

```{r}
mod_igp_res$par_summary
```
